---
title: "a3_task4_yvonne_dinh"
author: "Yvonne Dinh"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(kableExtra)
```

```{r}
#read in data
urchins=read.csv(here("data","urchins.csv")) %>% 
  clean_names()
```

##Comparing the mean size of purple urchins in control or annual sections of Naples Reef in January 2010
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#wrangle data to get purple urchin data for Naples in Jan 2010
naples=urchins %>% 
  filter(site=="NAPL") %>% 
  mutate(date = mdy(date)) %>% #convert the date to class "Date" 
  mutate(year= year(date)) %>% 
  mutate(month=month(date)) %>% 
  filter(year=="2010") %>% 
  filter(month=="1") %>% 
  filter(common_name=="Purple Urchin")

#histogram of the data!
ggplot(naples, aes(x=size))+
  geom_histogram()+
  labs(x="Urchin Diameter, cm", title= "Histogram of Purple Urchin Sizes in Naples Reef")+
  facet_wrap(~treatment)

#qqplot of the data!
ggplot(naples, aes(sample=size))+
  geom_qq()+
  facet_wrap(~treatment)
```

#Write 1 - 2 sentences describing your main takeaways about the distribution of urchin sizes for each group.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
#descriptive statistics table for both groups
naples_table=naples %>% 
  group_by(treatment) %>% 
  summarize(naples_mean = mean(size),
            naples_median = median(size),
            naples_sd=sd(size),
            naples_n=n())

naples_table

#preform a t-test to compare means between the two groups
annual_purple=naples %>% 
  filter(treatment == "ANNUAL") %>% 
  pull(size)

control_purple=naples %>% 
  filter(treatment == "CONTROL") %>% 
  pull(size)

t.test(x=annual_purple, control_purple)
```

# In text below your t-test code and output, write a comprehensive statement (2 - 3 sentences) comparing the means that provides meaningful discussion about the means difference beyond (but also including) the p-value, possibly including the actual means and standard deviations for each group, and the actual difference and / or % difference in means.

# In a final sentence, write the meaning of the p-value that you found in the t-test above. 


