---
title: "a3_task4_yvonne_dinh"
author: "Yvonne Dinh"
date: "11/1/2020"
output: html_document
---

```{r, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(kableExtra)
```

```{r, echo=TRUE, message=FALSE, warning=FALSE}
#read in data
urchins=read.csv(here("data","urchins.csv")) %>% 
  clean_names()
```

##Comparing the mean size of purple urchins in control or annual sections of Naples Reef in January 2010
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#wrangle data to get purple urchin data for Naples in Jan 2010
naples=urchins %>% 
  filter(site=="NAPL") %>% 
  mutate(date = mdy(date)) %>% #convert the date to class "Date" 
  mutate(year= year(date)) %>% 
  mutate(month=month(date)) %>% 
  filter(year=="2010") %>% 
  filter(month=="1") %>% 
  filter(common_name=="Purple Urchin")

#histogram of the data!
ggplot(naples, aes(x=size))+
  geom_histogram()+
  labs(x="Urchin Diameter, cm", title= "Histogram of Purple Urchin Sizes in Naples Reef")+
  facet_wrap(~treatment)

#qqplot of the data!
ggplot(naples, aes(sample=size))+
  geom_qq()+
  facet_wrap(~treatment)
```

Considering the histogram and qqplot above, the Annual group of Purple Urchins has a slightly larger distribution than the Control group. The spread is a bit wider in the Annual group. 


```{r, echo=TRUE, message=FALSE, warning=FALSE}
#descriptive statistics table for both groups
naples_table=naples %>% 
  group_by(treatment) %>% 
  summarize(naples_mean = mean(size),
            naples_median = median(size),
            naples_sd=sd(size),
            naples_n=n())
naples_table

#actual and percent difference calculations
actual_diff=(abs(3.91667-3.725806))
percent_diff=actual_diff/((4.469797+6.776489)/2)*100

#preform a t-test to compare means between the two groups
annual_purple=naples %>% 
  filter(treatment == "ANNUAL") %>% 
  pull(size)

control_purple=naples %>% 
  filter(treatment == "CONTROL") %>% 
  pull(size)

t.test(x=annual_purple, control_purple)
```
The mean sizes for the Annual and Control group are 3.92 cm and 3.73 cm, respectively, giving a mean difference of 0.16 cm. The percent difference in means is 3.39%. The standard deviation of the Annual and Control values are 0.966 and 0.867, respectively, indicating that the Annual group has slightly higher variance than the Control group. 

The p-value for the t-test above is 0.2536, meaning that there is a 25.36% probability that the two sample means are at least as different as what was observed (or 3.92-3.76 = 0.16cm). 

