---
title: "a3_task3"
author: "Yvonne Dinh"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
```


## Section 1: Size comparisons, purple and red urchins
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#read in data
urchins=read.csv(here("data","urchins.csv")) %>% 
  clean_names()

#create a faceted histogram of sea urchin sizes by common name
ggplot(urchins, aes(x=size))+
  geom_histogram()+
  facet_wrap(~common_name)+
  labs(x="Urchin Diameter (cm)")

#create a boxplot with common name oon x-axis and size on y-axis
ggplot(urchins, aes(x=common_name, y=size))+
  geom_boxplot()+
  labs(x="Common Name", y="Urchin Diameter (cm)")

#create a summary table that contains mean, median, standard deviation, and sample size for purple and red urchins
urchin_stats=urchins %>%
  group_by(common_name) %>%
  summarize(mean_urch=mean(size),
            med_urch=median(size),
            sd_urch=sd(size), #standard deviation
            n_urch=n()) #sample size
urchin_stats
actual_diff=(abs(4.469797-6.776489))
percent_diff=(abs(4.469797-6.776489))/((4.469797+6.776489)/2)*100
```

Looking at the histograms, the Red Urchins are more evenly distributed and have a wider distribution than the purple urchins. Most Purple Urchins diameters fall below 8 cm with a central tendency around 5 cm and most Red Urchins have a central tendency around 7 cm. The actual difference between the two species is 2.307 cm and the percent difference between the means of the two species is 41.021%

## Exploring purple urchin counts by site and year
```{r, echo=TRUE, message=FALSE, warning=FALSE}
#subset urchin data to only purple urchins
purple_urchins=urchins %>% 
  filter(common_name=="Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% #convert the date to class "Date" 
  mutate(year= year(date)) %>% #create a new column that contains year 
  mutate(site_full = case_when( #create a new column with full site names
    site=="NAPL" ~ "Naples Reef",
    site=="AQUE" ~ "Arroyo Quernardo Reef",
    site=="CARP" ~ "Carpinteria Reef",
    site=="MOHK" ~ "Mohawk Reef",
    site=="IVEE" ~ "Isla Vista Reef"))

#create a summary table containing the total number of purple urchins observed by site and year
purp_urch_site_year <- purple_urchins %>% 
  group_by(site, year) %>% 
  summarize(count = n())
purp_urch_site_year

#create an exploratory graph of counts of purple urchins recorded at each site
ggplot(purp_urch_site_year, aes(x=year, y=count, color=site))+
  geom_line()+
  scale_x_continuous(name = " ", breaks = c(2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020))+
  labs(title="Purple Urchin counts for each site")

        #scale_x_continuous
#what's up w the year decimals
```

